### 2021-06-09 paper/report.tex
Updates from Zoom meeting.

In our model, susceptibility to contracting the SARS-CoV-2 virus is dependent on age. A
possible infection progresses as shown in Figure~\ref{fig:disease_progression}. We
differentiate between an initial period of infection without being infectious or showing
symptoms, being infectious (presymptomatic or asymptomatic), showing symptoms, requiring
intensive care, and recovery or death. The probabilities of transitioning between these
states depend on age; their duration is random within intervals calibrated to medical
literature (for a detailed description see Section~\ref{sec:medical_params}).
Conditional on the type of contact, infectiousness independent of ages
\citep{Jones2021}.\comment[id=HM]{Seems broadly standard, can we cite someone?}

@@

We include two types of tests. Polymerase chain reaction (PCR) tests directly reveal
whether an individual is infected or not. PCR tests require some days to be processed
and there are always aggregate capacity constraints. In contrast, rapid antigen tests
yield immediate results. Specificity and sensitivity of these tests is set according to
data analysed in \cite{Bruemmer2021, Smith2021}; sensitivity depends on the timing of
the test relative to the start of infectiousness. Figure~\ref{fig:antigen_tests} shows
our model for rapid test demand. Schools may require students to be tested regularly.
Rapid tests may be offered by employers for on-site workers. Individuals may demand
tests for private reasons, which include having plans to meet other people\footnote{A
    positive test will make them reduce their contacts; this is why tests impact the actual
    contacts in Figure~\ref{fig:model-description}.}, showing symptoms of CoViD-19, and
because a household member tested positively for the virus. We endow each agent with an
individual compliance parameter. This parameter determines whether she takes up rapid
tests offered by employers or follows up on private reasons. Thresholds differ across
reasons triggering rapid test demand; they are lowest for infected household members and
highest for meeting other people.\comment[id=HM]{True?}\comment[id=K]{FALSE. We only have one private demand threshold (over time) which is the same for own symptoms and no PCR test, household member with symptoms or positive test and for having a weekly meeting with friends.}.

Modelling a population of agents according to actual demographic characteristics means
that we can use a wide array of data to identify and estimate the model's many
parameters.\footnote{See section S.X of the supplementary materials for an
    overview.}\comment[id=HM]{Need to make a list including symbols, maybe you can fetch the
parameters from the model and then we brainstorm re notation?} Mobility data is used to
model the reductions in work contacts. School and daycare policies are incorporated
directly from official directives. Infection rates by age and geographical region are
estimated to match to officially recorded numbers; so is the prevalence of virus
strains. The model yields total infections, but only a fraction of those will be
officially recorded. We model the fraction of known cases as depicted in
Figure~\ref{fig:model_for_official_cases}.\comment[id=HM]{Add a sentence.}
\footnote{EQUATIONS FOR THE FOURTH PICTURE (find names!):

@@

We apply this model to Germany. In March and April 2020, the country broke the first wave
of the pandemic fairly quickly. Between mid-May and mid-September, daily new infections
were below 20 per Million and day.\comment[id=HM]{Cite our world in data} We model the
period mid-September 2020 to the end of May 2021. We pick the starting date for two
reasons. First, we do not include the first wave because the environment was very
different (e.g., aggregate PCR test capacity was much lower and we would require a very
different model for calculating the share of known cases) Second, a large fraction cases
during summer of 2020 were traced to international travel
\citep{KochInstitut2021,Hodcroft2021} but the precise number is difficult to model.

Figure~\ref{fig:pandemic_drivers_model_fit} describes the evolution of the pandemic and
of its drivers. The black line in Figure~\ref{fig:aggregated_fit} shows officially
recorded cases; the \replaced[id=HM]{color}{blue?} line in
Figure~\ref{fig:stringency_index} the Oxford Response Stringency Index \citep{Hale2020},
which tracks the tightness of non-pharmaceutical interventions. We transform the index
so that lower values represent higher levels of restrictions. A value of zero means all
measures incorporated in the index are turned on. The value 1 represents the situation
in mid-September, with restrictions on gatherings and public events, masking
requirements, but open schools and workplaces (the raw value of the index at that point
is 49.5). In the seven weeks between mid September and early November, cases increased
by a factor of 10. Restrictions were somewhat tightened in mid-October and again in
early November. New infections remained constant throughout November, before rising
again in December, which prompted the most stringent lockdown to this date. Schools and
daycare centers were closed again, so were customer-facing businesses except for grocery
and drug stores. From the peak of the second wave just before Christmas until the trough
in mid-February, newly detected cases decreased by almost three quarters. The third wave
in the spring of 2021 is associated with the B.1.1.7 strain, which became dominant in
March. See Figure~\ref{fig:share_b117}. In early March, some NPIs were being relaxed;
e.g., hairdressers and home improvement stores were allowed to open again to the public.

@@

Our model is able to track these developments very well. The blue line in
Figure~\ref{fig:aggregated_fit} shows our model's predictions are very close to
officially recorded cases in the aggregate. This is also true for infections by age and
geographical region, which are shown in the supplementary materials
(Figures~\ref{fig:age_group_fit} and \ref{fig:state_fit}, respectively). We can
disentangle various mechanisms due to the distinct temporal variation in the drivers of
the pandemic. Next to the stringency index, the three lines in
Figure~\ref{fig:stringency_index} summarize how contact reductions, increased hygiene
regulations, and seasonality evolved since early September for each of the three broad
contact networks. For example, a value of 0.75 for the work multiplier means that if the
environment was the same as in September (levels of infection rates, no rapid tests or
vaccinations, only the wildtype virus present), infections at the workplace would be
reduced by 25\%. This reduction is the product of the effect of contact reductions,
increased hygiene regulations, and seasonality. Along with the levels of infections,
these measures completely determine the spread of SARS-CoV-2 in
2020.\comment[id=HM]{True?} See Appendix~\comment[id=HM]{Reference}
\comment[id=K]{We have not done the decomposition for 2020 yet. Is on my todo list}
\comment[id=HM]{I guess my question was whether I was missing any factors.}
for their individual contributions. Two things are particularly interesting. First, all
lines broadly follow the stringency index and they would do so even more if we left out
seasonality and school vacations (roughly the last two weeks of October, two weeks each
around Christmas and Easter, and some days in late May). Second, the most stringent
regulations are associated with the period of strong decreases in new infection between
late December 2020 and mid-February 2021. The measures were not enough, however, to stop
the B.1.1.7 variant from spreading in the subsequent period. The steep drop in recorded
cases during May 2021 is associated with at least weekly rapid tests to around
42~percent of the population, a vaccination rate that rose from 28\% to 43\%, and mostly
seasonality impacting a fall in the relative infectiousness of contacts outside of work
and school.

Figure~\ref{fig:2021_scenarios_broad} consider the relative effects of rapid tests,
vaccinations, and of seasonality during 2021, assuming NPIs to have evolved the same way
as in the baseline scenario. Figure~\ref{fig:2021_scenarios_recorded} shows the model
fit (the bold blue line\comment[id=HM]{Make sure this is the case}, same as in
Figure~\ref{fig:aggregated_fit}), a scenario without any of the three factors (x line),
and three scenarios turning these factors off one by one.
Figure~\ref{fig:2021_scenarios_newly_infected} does the same for total infections in the
model. Figure~\ref{fig:2021_scenarios_decomposition} employs Shapley values to decompose
the difference in total infections between the scenario without any of the three factors
and our main specification.

>>>>

In our model, susceptibility to contracting the SARS-CoV-2 virus is dependent on
age. A possible infection progresses as shown in
Figure~\ref{fig:disease_progression}. We differentiate between an initial period
of infection without being infectious or showing symptoms, being infectious
(presymptomatic or asymptomatic), showing symptoms, requiring intensive care,
and recovery or death.\comment[id=HM]{Seems broadly standard, can we cite
someone?} The probabilities of transitioning between these states depend on age;
their duration is random within intervals calibrated to medical literature (for
a detailed description see Section~\ref{sec:medical_params}). Conditional on the
type of contact, infectiousness is independent of age \citep{Jones2021}.

@@

We include two types of tests. Polymerase chain reaction (PCR) tests directly reveal
whether an individual is infected or not. PCR tests require some days to be processed
and there are always aggregate capacity constraints. In contrast, rapid antigen tests
yield immediate results. Specificity and sensitivity of these tests is set according to
data analysed in \cite{Bruemmer2021, Smith2021}; sensitivity depends on the timing of
the test relative to the start of infectiousness. Figure~\ref{fig:antigen_tests} shows
our model for rapid test demand. Schools may require students to be tested regularly.
Rapid tests may be offered by employers for on-site workers. Individuals may demand
tests for private reasons, which include having plans to meet other people\footnote{A
    positive test will make them reduce their contacts; this is why tests impact the actual
    contacts in Figure~\ref{fig:model-description}.}, showing symptoms of CoViD-19, and
because a household member tested positively for the virus. We endow each agent with an
individual compliance parameter. This parameter determines whether she takes up rapid
tests offered by employers or follows up on private reasons. The thresholds are lower for
tests in a private setting than for tests at the workplace.\comment[id=HM]{True? @Janos}

Modelling a population of agents according to actual demographic characteristics means
that we can use a wide array of data to identify and estimate the model's many
parameters.\footnote{See section S.XXX of the supplementary materials for an
    overview.} Mobility data is used to
model the reductions in work contacts. School and daycare policies are incorporated
directly from official directives. Infection rates by age and geographical region are
estimated to match to officially recorded numbers; so is the prevalence of virus
strains. The model yields total infections, but only a fraction of those will be
officially recorded. We model the fraction of known cases as depicted in
Figure~\ref{fig:model_for_official_cases}.\comment[id=HM]{Add a sentence.}
\footnote{EQUATIONS FOR THE FOURTH PICTURE (find names!):

@@

We apply this model to Germany. In March and April 2020, the country broke the first wave
of the pandemic fairly quickly. Between mid-May and mid-September, daily new infections
were below 20 per Million and day.\comment[id=HM]{Cite our world in data/RKI} We model the
period mid-September 2020 to the end of May 2021. We pick the starting date for two
reasons. First, we do not include the first wave because the environment was very
different (e.g., aggregate PCR test capacity was much lower and we would require a very
different model for calculating the share of known cases) Second, a large fraction cases
during summer of 2020 were traced to international travel
\citep{KochInstitut2021,Hodcroft2021} but the precise number is difficult to model.

Figure~\ref{fig:pandemic_drivers_model_fit} describes the evolution of the
pandemic and of its drivers. The black line in Figure~\ref{fig:aggregated_fit}
shows officially recorded cases; the black line in
Figure~\ref{fig:stringency_index} the Oxford Response Stringency Index
\citep{Hale2020}, which tracks the tightness of non-pharmaceutical
interventions. We transform the index so that lower values represent higher
levels of restrictions. A value of zero means all measures incorporated in the
index are turned on. The value 1 represents the situation in mid-September, with
restrictions on gatherings and public events, masking requirements, but open
schools and workplaces (the raw value of the index at that point is 49.5). In
the seven weeks between mid September and early November, cases increased by a
factor of 10. Restrictions were somewhat tightened in mid-October and again in
early November. New infections remained constant throughout November, before
rising again in December, which prompted the most stringent lockdown to this
date. Schools and daycare centers were closed again, so were customer-facing
businesses except for grocery and drug stores. From the peak of the second wave
just before Christmas until the trough in mid-February, newly detected cases
decreased by almost three quarters. The third wave in the spring of 2021 is
associated with the B.1.1.7 strain, which became dominant in March. See
Figure~\ref{fig:share_b117}. In early March, some NPIs were being relaxed; e.g.,
hairdressers and home improvement stores were allowed to open again to the
public.

@@

Our model is able to track these developments very well. The blue line in
Figure~\ref{fig:aggregated_fit} shows our model's predictions are very close to
officially recorded cases in the aggregate. This is also true for infections by
age and geographical region, which are shown in the supplementary materials
(Figures~\ref{fig:age_group_fit} and \ref{fig:state_fit}, respectively). We can
disentangle various mechanisms due to the distinct temporal variation in the
drivers of the pandemic. Next to the stringency index, the three lines in
Figure~\ref{fig:stringency_index} summarize how contact reductions, increased
hygiene regulations, and seasonality evolved since early September for each of
the three broad contact networks. For example, a value of 0.75 for the work
multiplier means that if the environment was the same as in September (levels of
infection rates, no rapid tests or vaccinations, only the wildtype virus
present), infections at the workplace would be reduced by 25\%. This reduction
is the product of the effect of contact reductions, increased hygiene
regulations, and seasonality. Along with the levels of infections, these
measures determine the spread of SARS-CoV-2 in 2020. See
Appendix~\comment[id=HM]{Reference} for separate plots of the three factors. Two
aspects are particularly interesting. First, all lines broadly follow the
stringency index and they would do so even more if we left out seasonality and
school vacations (roughly the last two weeks of October, two weeks each around
Christmas and Easter, and some days in late May). Second, the most stringent
regulations are associated with the period of strong decreases in new infection
between late December 2020 and mid-February 2021. The measures were not enough,
however, to stop the B.1.1.7 variant from spreading in the subsequent period.
The steep drop in recorded cases during May 2021 is associated with at least
weekly rapid tests to around 42~percent of the population, a vaccination rate
that rose from 28\% to 43\%, and mostly seasonality impacting a fall in the
relative infectiousness of contacts outside of work and school.

Figure~\ref{fig:2021_scenarios_broad} consider the relative effects of rapid
tests, vaccinations, and of seasonality during 2021, assuming NPIs to have
evolved the same way as in the baseline scenario.
Figure~\ref{fig:2021_scenarios_recorded} shows the model fit (the blue line,
same as in Figure~\ref{fig:aggregated_fit}), a scenario without any of the three
factors (red line), and three scenarios turning these factors off one by one.
Figure~\ref{fig:2021_scenarios_newly_infected} does the same for total
infections in the model. Figure~\ref{fig:2021_scenarios_decomposition} employs
Shapley values to decompose the difference in total infections between the
scenario without any of the three factors and our main specification.
